---
title: "ML1HW2"
author: "Adil Hayat"
date: "26 January 2017"
output:
  pdf_document: default
  html_document: default
---
## HW2 Part1
### Question 1
```{r}
suppressPackageStartupMessages(library(tidyverse))
library(stringr)
library(bestglm)

bonddata <-  read.table("http://www.stat.cmu.edu/~cschafer/MSCF/bonddata.txt",sep=",", header=T)
bonddata <- tbl_df(bonddata)
colnames(bonddata)

# Assumed that the columns with less than 10 distinct values as categorical 
# Extract the predictors from time steps 2 and prior 
prior_colnames<- sapply(names(bonddata), function(x) { str_match_all(x, "[0-9]+") %>% 
    unlist %>% unique %>% as.numeric >=2 })
prior_colnames <- names(prior_colnames[prior_colnames==TRUE]) %>% na.omit %>% unique

# columns to remove, removing the columsn bond_id and id as well 
col_remove <- c(prior_colnames,"id","bond_id")

bonddata_sel <- select(bonddata,-one_of(col_remove))

Xyframe <- data.frame(cbind(bonddata_sel[,(2:ncol(bonddata_sel))],bonddata_sel[,1]))
bestmod <- bestglm(Xyframe, IC="AIC")

summary(bestmod$BestModel)
```
### The categorical predictors trade_type and trade_type1 show up in the final model.


### Question 2
```{r}

fullmod <- lm(trade_price~.,data = bonddata_sel)
levs <- hatvalues(fullmod)

# Calculate PRESS
PRESS <- sum((fullmod$residuals/(1-levs))^2)

# comparing the PRESS of the model using the model estimated using AIC
levsbestmod <- hatvalues(bestmod$BestModel)

# Calculate PRESS
PRESSbestmod <- sum((bestmod$BestModel$residuals/(1-levsbestmod))^2)

# print the values
cat("The PRESS of the full model=",PRESS,"and the AIC fitted best model=",
    PRESSbestmod,"\n")
```
### Question 3
```{r}
# Using the Cook's distance
cookd = as.numeric(cooks.distance(bestmod$BestModel))
numcoeff <- length(bestmod$BestModel$coefficients)
sort(pf(cookd, numcoeff, nrow(bonddata_sel) - numcoeff),decreasing=TRUE)[1:5]
```
### The observations above 50th percentile are a cause of concern.
### The largest value of Cook's distance is 2% so there are no influentail observations.

### Question 4
```{r}
library(MASS)

holdrlm = rlm(as.formula((paste("trade_price",paste(names(bestmod$BestModel$coefficients[-1]),
                                                    collapse="+"),sep="~"))),data = bonddata_sel)

dat <- data.frame(holdrlm$fitted.values,holdrlm$residuals)
names(dat) <- c("Fitted_Values","Residuals")
ggplot(dat, aes(x=Fitted_Values, y=Residuals)) +
    geom_point(shape=1)
qqnorm(dat$Residuals);qqline(dat$Residuals, col=2)

```
### The residuals have heavy tails compared to a Normal Distibution

### Question 5
```{r}
library(car)
boxcox(lm(trade_price~.,data = bonddata_sel))
```
### This gives us the value of $\lambda$=1. So, no transformation of the data is required. 

\newpage
# Adil Hayat
## HW2Part2
### Questoin 1
```{r }
trainset <- readRDS("traindata.rds")
trainset <- tbl_df(trainset)
fullrow = rep(FALSE,nrow(trainset))
# commenting this code for efficiency
# for(i in 1:nrow(trainset))
# {
#   fullrow[i] = !any(is.na(trainset[i,29:147]))
# }

fullrow <- apply(trainset, 1, function(x){!any(is.na(x[29:147]))})
```
fullrow is a boolean vector that is TRUE only if all the values in a row are available specifically from columns from 29 to 147(are not NAs)   

### Questoin 2
```{r}
varnames <- c(paste("Ret_", 2:120, sep=""))
fullform = as.formula(paste("Ret_PlusOne ~ ",paste(varnames,collapse="+")))
```
fullform contains the formula for the regression with Ret_PlusOne as the response variable and the variables Ret_2 to Ret_120 as the predictors.

### Questoin 3
```{r}
filt_trainset <- trainset[fullrow,]
fullmod <- lm(fullform,data=filt_trainset)
summary(fullmod)
```

### Question 4
Using the step function to find the best model
```{r}
finalmod = step(fullmod, direction="both", trace = FALSE)
```

### Question 5
plotting the residuals and QQPlot
```{r}
dat= data.frame(finalmod$fitted.values,finalmod$residuals)
names(dat) <- c("Fitted_Values","Residuals")
ggplot(dat, aes(x=Fitted_Values, y=Residuals)) +
    geom_point(shape=1)
qqnorm(dat$Residuals);qqline(dat$Residuals, col=2)
```
### Question 6
Cook's Distance Calculation
```{r}
numpred = length(finalmod$coefficients)
cookd = as.numeric(cooks.distance(finalmod))
sort(pf(cookd,numpred,nrow(trainset)-numpred),decreasing=TRUE)[1:5]
```
# There are two data points with the pencentile of Cook's distance of 100% and 99.99%. So, these 2 data points are very influential.
